
\documentclass[a4paper,12pt]{report} 

\usepackage[top = 3.4cm, left = 2.4cm, right = 2.4cm, bottom = 2.6cm]{geometry}
\usepackage{multicol}

\usepackage{tikz}
\usepackage{draftwatermark}
\SetWatermarkText{
\begin{tikzpicture}
\draw[fill] (-1.4142, 0) -- (0, -1) -- (1.4142, 0) -- (0, 1) -- cycle;
\end{tikzpicture}
}
\SetWatermarkHorCenter{2cm}
\SetWatermarkVerCenter{2cm}
\SetWatermarkColor[rgb]{0.35,0.65,0.9}

\usepackage{hyperref}
\hypersetup{
	colorlinks=true,
	urlcolor=black
}

\usepackage{natbib}
%\usepackage{apacite}
%\bibliographystyle{apacite}
\usepackage{bibentry}
\nobibliography*

\begin{document} 

\begin{multicols}{2}
[
\begin{center}
{\Large \textbf{Lyron Winderbaum}} \\ \vspace{4mm}
\end{center}
]

\begin{flushleft}
\begin{tabular}{ll}
%\textbf{Details} & \\ 
mobile: & (+61) 431 955 152 \\
email: & \href{mailto:armadilloa16@gmail.com}{armadilloa16@gmail.com} \\
\end{tabular}
\end{flushleft}

\begin{flushright}
\begin{tabular}{l}
%\textbf{Address} \\
%3 Hallett Avenue \\  
address: Tranmere, South Australia \\
%Australia \\
website: \href{https://armadilloa16.github.io/eportfolio/}{{\footnotesize armadilloa16.github.io/eportfolio/}} \\
{}
\end{tabular}
\end{flushright}
\end{multicols}                                  

\hrule
\vspace{0.4cm}

Highly qualified \textbf{Mathematics and Science Educator} with extensive experience in university education and one-on-one mathematics tutoring. Both academic research, industry, and teaching experience in mathematics, statistics and science. 

\vspace{0.4cm}
\hrule

\section*{Education and Credentials}
(The University of Adelaide)

\subsubsection*{Master of Teaching \hfill Jan 2018 - Completed: July 2019}
Thesis Title: ``MathsStart and MathsTrack: Mathematics Bridging at the University of Adelaide.'' \\
Supervisors: Dr. Igusti Darmawan, and collaboration with Mr. Nicholas Crouch.

Degree involved research, coursework, and placement components. Research component focused on mathematics education, maths anxiety, and the secondary-tertiary education transition including a specific focus on mathematics bridging courses. Coursework components focussed on inclusive education and both learning theory and practical approaches to supporting students. Placement component involved 13 weeks of teaching experience in high schools, teaching mathematics and science from years 8-12.


\subsubsection*{Doctor of Philosophy in Statistics \hfill Conferred: Dec 2016} \vspace{-6pt}
Thesis: "Statistical Treatment of Proteomic Imaging Mass Spectrometry Data", supervised by Professor Inge Koch and Professor Peter Hoffmann.

Developed, implemented and aided interpretation of results from processing and analysis of data from imaging mass spectrometry experiments. Methods used included principal component analysis, k-means clustering, and discriminant analysis, and broadly fit under the umbrella of machine learning. Datasets related to a variety of contexts but primarily biomedical applications to gynacological cancer treatement. The work done towards this degree involved a significant component of software development in MATLAB and R, and writing in \LaTeX using  \href{https://yihui.name/knitr/}{knitr}.

\subsubsection*{Honours in Statistics \hfill Conferred: Dec 2011} \vspace{-6pt}
Thesis: "Multivariate Analysis of Trace Elements in Pyrite", supervised by Associate Professor Andrew Metcalfe, Professor Nigel Cook, and Dr. Cristiana Ciobanu.

Developed methods through collaboration with geochemistry experts to investigate the relationships of interest to them. Methods such as hierarchical agglomerative cluster analysis and bootstrapping were adapted to produce results that were meaningful to the geologists.


\subsubsection*{Bachelor of Science (Chemistry) \hfill Conferred: Mar 2011}

\subsubsection*{Bachelor of Computer and Mathematical Sciences \hfill Conferred: Dec 2010 \\ (Pure Mathematics and Statistics)}

\vspace{0.4cm}
\hrule
\vspace{0.6cm}

\section*{Additional Qualifications}

\subsubsection*{Understanding Aboriginality Webinar \hfill attended: 10 May 2018}

\subsubsection*{Keeping Safe: Child Protection Curriculum Training \hfill completed: 19 Apr 2018}

\subsubsection*{Youth Mental Health First Aid \hfill accredited: 5 Mar 2018}

%\subsubsection*{Literacy and Numeracy Test for Initial Teacher Education Students \hfill standard achieved: 3 Mar 2018}
%
%\subsubsection*{Responding to Abuse and Neglect Course \hfill completed: 22 Feb 2018}
%
%\subsubsection*{Department for Communities and Social Inclusion Child-Related Employment Screening \hfill cleared: 16 Jan 2018}

\subsubsection*{Basic Emergency Life Support (HLTAID002) \hfill accredited: 16 Jun 2017}






\vspace{0.4cm}
\hrule

\vspace{1cm}
\section*{Professional Experience}

\subsubsection*{Statistician  \hfill Mar 2019 - Present} \vspace{-6pt}
(LBT Innovations) \\[4pt]

Developing and validating in-vitro diagnostic medical devices, including planning, performing, 
and reporting the analysis of both analytical and clinical studies of performance.

\subsubsection*{Postdoctoral Research Associate  \hfill Jan 2017 - Dec 2017} \vspace{-6pt}
(Future Industries Institute, University of South Australia) \\[4pt]

Extensive collaboration with experts in different fields and adapting statistical
analyses to fullfil their needs. Led and mentored colleagues to implement
improved data storage and reporting techniques to ensure a high standard of transparency
and accountability. Used high level analytical, research, and communication skills to develop, 
implement, and report on various multivariate statistical data analyses.

\subsubsection*{Research Associate  \hfill Aug 2015 - Dec 2016} \vspace{-6pt}
(Adelaide Proteomics Centre, The University of Adelaide) \\[4pt]

Developed and implemented software solutions to improve work efficiency
and provide quality assurance to customers. Collaboration with colleagues and experts
in different fields, particularly biomedical research, and adaption statistical analyses to
different scenarios. Role required high level analytical, research, and communication skills
were required in order to develop, implement, and report on various multivariate statistical
data analyses.

\subsubsection*{Tutor \hfill Mar 2012 - Mar 2019} \vspace{-6pt}
(Maths Learning Centre, The University of Adelaide) \\[4pt]

\subsubsection*{Private Tutor \hfill Mar 2018 - Present} \vspace{-6pt}
(sub-contracted through Tutoring for Excellence) \\[4pt]

Present complex information to people from a diverse range of backgrounds. Requires a high level of interpersonal skills and the ability to adapt material to different cultures and learning styles. Consulted research students from accross every dicipline in the university on statistical analyses needed for their projects. Delivered the statistics component of third year medical students research methods component. Communicated mathematics and statistics to a wide variety of learners, from year 7 school students to higher degree research students, with adjustments to their individual perspectives and needs.

\subsubsection*{Tutor \hfill Mar 2012 - July 2015} \vspace{-6pt}
(School of Mathematical Sciences, The University of Adelaide) \\[4pt]

Experience teaching tutorials, supervising computer practicals, and marking assignments and exams under time-constraints.

\subsubsection*{Storeperson \hfill Feb 2007 - Feb 2012} \vspace{-6pt}
(Coles Supermarkets) \\[4pt]

Retail and customer service experience.


\vspace{1cm}
\hrule

\section*{Academic Awards}

\subsubsection*{Australian Postgraduate Award Scholarship \hfill Conferred: 2012}

\subsubsection*{The Applied Probability Trust Prize \hfill Conferred: 2011}

\subsubsection*{Australian Bureau of Statistics Scholarship \hfill Conferred: 2008}

\vspace{0.4cm}
\hrule
\vspace{0.6cm}

\subsection*{Programming and Computer Skills}          

\begin{itemize}
	\item Extensive experience with MATLAB, \LaTeX, and R.
	\item Familiarity with Python, Java, Golang, C\#, Ruby, Perl, Git, and other software such as SAS/ SPSS. 
\end{itemize}


%\hrule

\subsection*{Publications}          

\begin{itemize}
	\item \bibentry{Gustafsson2018}
	\item \bibentry{Gustafsson2018a}
	\item \bibentry{Arentz2017}
	\item \bibentry{Mittal2017}
	\item \bibentry{Zammit2016}
	\item \bibentry{Zhang2016}
	\item \bibentry{Mittal2016}
	\item \bibentry{Winderbaum2016}
	\item \bibentry{Winderbaum2015}
	\item \bibentry{Gustafsson2015}
	\item \bibentry{Winderbaum2012}
\end{itemize}


\hrule


\subsection*{Presentations}          

\begin{itemize}
	\item Oral Presentation, Mathematical Association of South Australia Annual Conference, 2019
	\item Poster Presentation, Australian Proteomics Society Annual Lorne Symposium, 2015
	\item Oral Presentation, Australian Statistican Conference/ Institute of Mathematical Statistics Annual Meeting, Sydney, 2014
	\item Oral Presentation, Young Statisticians Conference, Brisbane, 2013
\end{itemize}


\hrule


%\begin{multicols}{2}
%[
%\subsection*{References}
%]
%
%\begin{flushleft}
%\textbf{John Doe} \\[2pt]
%\begin{tabular}{ll}
%%Position: & Executive Director, \\
%%	    & Choose Maths \\
%Relation: & Current Employer, \\
%               & Colleague \\
%%Phone:    & +61 3 8344 9859 \\
%Phone:    & +61 123 456 789 \\
%Email:      & \href{mailto:john.doe@gmail.com}{john.doe@gmail.com} \\
%\end{tabular}
%\end{flushleft}
%
%\begin{flushleft}
%\textbf{John Doe} \\[2pt]
%\begin{tabular}{ll}
%%Position: & Executive Director, \\
%%	    & Choose Maths \\
%Relation: & Current Employer, \\
%               & Colleague \\
%%Phone:    & +61 3 8344 9859 \\
%Phone:    & +61 123 456 789 \\
%Email:      & \href{mailto:john.doe@gmail.com}{john.doe@gmail.com} \\
%\end{tabular}
%\end{flushleft}
%
%\end{multicols}


%\bibliographystyle{plain}
\bibliographystyle{plainnat}
\nobibliography{publications}

\end{document}
